---
title: "Parameter Estimation with Priors"
author: "Tripp Bishop"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(latex2exp)
theme_set(theme_minimal())
```

We'll consider the weak prior first.

```{r}
x <- seq(0,1,0.001)
y_prior <- dbeta(x,7,3)
y_data <- dbeta(x,25,17)

tibble(x=x,y_prior=y_prior, y_data=y_data) |> 
  ggplot(aes(x=x)) +
  geom_line(aes(y=y_prior), colour="red", linewidth=1) +
  geom_line(aes(y=y_data), colour="blue", linewidth=1) +
  labs(
    x="Quantile",
    y="Likelihood",
    title=TeX("PDF for Prior (red) and Prior + Data (blue)")
  )
```
Plot the CDF to get some rough estimates for the upper and lower bounds of the confidence interval.

```{r}
q <- seq(0,1,0.001)
y_cdf <- pbeta(q, 25, 17)
p_lower <- 0.025
p_upper <- 0.975
tibble(x=x,y=y_cdf) |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(colour="red", linewidth=1) +
  geom_hline(yintercept = p_lower, linetype="dashed") +
  geom_hline(yintercept = p_upper, linetype="dashed") +
  labs(
    x="Quantiles",
    y="Cumulative Probability",
    title=TeX("CDF for $\\beta(25, 17)$ distribution")
  )
```
The confidence interval ranges between `r qbeta(0.025, 25,17)` and `r qbeta(0.975, 25,17)`.

Now the strong fair prior.

```{r}
x <- seq(0,1,0.001)
y_prior <- dbeta(x,1000,1000)
y_data <- dbeta(x,1018,1014)

tibble(x=x,y_prior=y_prior, y_data=y_data) |> 
  ggplot(aes(x=x)) +
  geom_line(aes(y=y_prior), colour="red", linewidth=1) +
  geom_line(aes(y=y_data), colour="blue", linewidth=1) +
  labs(
    x="Quantile",
    y="Likelihood",
    title=TeX("PDF for Prior (red) and Prior + Data (blue)")
  )
```
The strong prior completely overwhelms the scant data we have.

```{r}
q <- seq(0,1,0.001)
y_cdf <- pbeta(q, 1018, 1014)
p_lower <- 0.025
p_upper <- 0.975
tibble(x=x,y=y_cdf) |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(colour="red", linewidth=1) +
  geom_hline(yintercept = p_lower, linetype="dashed") +
  geom_hline(yintercept = p_upper, linetype="dashed") +
  labs(
    x="Quantiles",
    y="Cumulative Probability",
    title=TeX("CDF for $\\beta(1018, 1014)$ distribution")
  )
```
Consequently, we can see that the 95% confidence interval is quite narrow.

The confidence interval ranges between `r qbeta(0.025, 1018,1014)` and `r qbeta(0.975, 1018,1014)`.

And now, let's look at the modestly biased prior.

```{r}
x <- seq(0,1,0.001)
y_prior <- dbeta(x,70,30)
y_data <- dbeta(x,88,44)

tibble(x=x,y_prior=y_prior, y_data=y_data) |> 
  ggplot(aes(x=x)) +
  geom_line(aes(y=y_prior), colour="red", linewidth=1) +
  geom_line(aes(y=y_data), colour="blue", linewidth=1) +
  labs(
    x="Quantile",
    y="Likelihood",
    title=TeX("PDF for Prior (red) and Prior + Data (blue)")
  )
```
The two PDFs are close, but we can see that the data have pulled the distribution down closer to even odds. Still, the data aren't overwhelming the prior by any stretch.

```{r}
q <- seq(0,1,0.001)
y_cdf <- pbeta(q, 88, 44)
p_lower <- 0.025
p_upper <- 0.975
tibble(x=x,y=y_cdf) |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(colour="red", linewidth=1) +
  geom_hline(yintercept = p_lower, linetype="dashed") +
  geom_hline(yintercept = p_upper, linetype="dashed") +
  labs(
    x="Quantiles",
    y="Cumulative Probability",
    title=TeX("CDF for $\\beta(88, 44)$ distribution")
  )
```
The range is not so tight here since we don't have a strong prior or a lot of data.

The confidence interval ranges between `r qbeta(0.025, 88,44)` and `r qbeta(0.975, 88,44)`.
