---
title: "The Normal Distribution"
format: html
theme: sandstone
---

```{r setup}
library(tidyverse)
library(latex2exp)
theme_set(theme_minimal())
```


## Notes

## Exercises Chapter 12

### Problem 1
What is the probability of observing a value five $\sigma$ greater than the mean or more?

We can solve this by using the standard normal distribution ($\mu=0$, $\sigma=1$) and using the `pnorm` function to determine the area under the PDF curve for $q\geq 5$.

```{r}
pnorm(5, lower.tail = FALSE)
```
As expected, the probability is very low.

### Problem 2
A fever is any temperature greater than 100.4 degrees Fahrenheit. Given the following measurements, what is the probability that the patient has a fever?

$$
100.0, 99.8, 101.0, 100.5, 99.7
$$
First, we compute the mean and standard deviation from which we can create a probability distribution. Then, we can determine how much probability lies to the right of 100.4 degrees given the distribution.

```{r}
temps <- c(100.0, 99.8, 101.0, 100.5, 99.7)
mean_temp <- mean(temps)
sd_temp <- sd(temps)

pnorm(100.4, mean=mean_temp, sd=sd_temp, lower.tail = FALSE)
```
There is a 35.6% chance that the patient has a fever based on the data available.

### Problem 3
Suppose in Chapter 11 we tried to measure the depth of a well by timing coin drops and got the following values:
$$
2.5, 3, 3.5, 4, 2
$$
The distance an object falls can be calculated (in meters) with the following formula:
$$
\text{distance} = 1/2 \times G \times \text{time}^2
$$
where $G$ is 9.8 m/s/s. What is the probability that the well is over 500 meters deep?

To answer this, we need to determine the mean and standard deviation and then make the assumption that the data are normally distributed (seems like big if, but that's what we're doing).

```{r}
times <- c(2.5, 3, 3.5, 4, 2)
time_mean <- mean(times)
time_sd <- sqrt(1/length(times)*sum((times - time_mean)^2))
```

This gives us the following distribution, $N(3,0.79)$. If the well were 500m deep, then the time to sound would be $1/2*9.8*t^2 = 500$. Solving for $t$ yields $t=\sqrt{2*500/9.8} \approx 10.1$. This value is many standard deviations above the mean, so we would expect this to be *highly* unlikely. Using the `pnorm` function, we get a probability of
```{r}
pnorm(10.1, time_mean, time_sd, lower.tail = FALSE)
```
which is effectively 0.

### Problem 4
What is the probability that there is no well (i.e the well is really 0 meters deep)? You'll notice that probability is higher than you might expect, given your observation that there *is* a well. There are two good explanations for this probability being higher than it should. The first is that the normal distribution is a poor model for our measurements: the second is that, when making up numbers for an example, I chose values that you likely wouldn't see in real life. Which is more likely to you?

```{r}
pnorm(0, time_mean, time_sd)
```
There really isn't a reason to mistrust the normal distribution assumption here, so we might consider if the data are bad.


```{r}
#| warning: false
x <- seq(from=-4, to=4, by=0.01)
y1 <- dnorm(x)
y2 <- dnorm(x, 0, 0.5)
y3 <- dnorm(x, 0, 2)
tibble(x=x,y1=y1, y2=y2, y3=y3) |> 
  pivot_longer(cols=-x, names_to = "y", values_to = "vals") |> 
  ggplot(aes(x=x,y=vals)) +
  geom_line(aes(colour=y, linetype=y), linewidth=1) +
  annotate("text", x=1.40, y=0.24, label=TeX("$\\sigma=1$")) +
  annotate("text", x=0.8, y=0.62, label=TeX("$\\sigma=0.5$")) +
  annotate("text", x=3, y=0.11, label=TeX("$\\sigma=2$")) +
  labs(
    x="x",
    y="y"
  ) +
  theme(
    legend.position = "none"
  )
```
```{r}
#| warning: false
df <- tibble(x=x,y=y1)
df |> 
  ggplot(aes(x=x,y=y)) +
  geom_line(linewidth=1) +
  geom_ribbon(data=subset(df, x>=-1 & x<=1), aes(x=x, ymin=0,ymax=y), alpha=0.4, fill="steelblue") +
  annotate("text", x=1.40, y=0.24, label=TeX("$\\sigma=1$")) +
  labs(
    x="x",
    y="y"
  ) +
  theme(
    legend.position = "none"
  )
```


